---
title: "Lab 1"
author: "Asha Yadav"
date: "October 7, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(error = TRUE)
```

# Loading a library
Run the code chunk below to load the tidyverse. If you get errors, you probably 
don't have it installed yet, so run `install.packages("tidyverse")` first.

```{r load-tidyverse, message = FALSE, warning = FALSE}
library(tidyverse)
```

Note that you should get some messages printed out below that state all the 
packages that have been loaded. This is important to pay attention to. It says
that both *ggplot2* and *dplyr* have been loaded, so you don't need explicit 
calls to those packages 

Many packages also include data. We'll focus here on the *mpg* dataset, which
comes from *ggplot2*. Run the following code to produce a scatterplot 

```{r first-plot, options}
ggplot(mpg, aes(x = displ, y = hwy)) +
  geom_point()
```

Look at the help documentation for `geom_point`. Note you'll have to run the 
code to access the help documentation within a code chunk, rather than in the
text areas. Try changing the color of the points to "gray70". 
```{r geom_point}
help("geom_point")
```
```{r gray 70}
ggplot(mpg, aes(x = displ, y = hwy)) + geom_point(color="gray 70")
```


In the code chunk below create a second scatterplot with the *mpg* dataset, this time with the number of cylinders displayed on the x-axis and the city miles per gallon on the y-axis. Also change the color of the points to "gray70".

```{r second-plot}
ggplot(mpg, aes(x = cyl, y = cty)) + geom_point(color="gray 70")
```

The *diamonds* dataset is also included with *ggplot2*. The code below will
show the first six rows of the dataset.

```{r head-diamonds}
head(diamonds)
```

Look at the help documentation and then use the code below to change the 
number of rows shown to 15.

```{r head-diamonds15}
help("head")
head(diamonds, n=15 )
```

Call a similar function that shows the last six rows of the dataset. (Note you may need to look at the same help documentation.)

```{r last-rows}
tail(diamonds)
```

Use any function you want to determine the dimensions of the dataset - i.e., the number of rows and columns.

```{r diamonds-dimensions}
dim(diamonds)
```

Use the pipe operator, `%>%`, along with `count`, to count the rows of each diamond color. 

```{r diamond-colors1}
diamonds %>%
  count(color) %>%
  head(15)
```

Write the code again, this time without the pipe operator

```{r diamond-colors2}
head(count(diamonds, color), 15)
```

Be sure to add your name to the top of the document (line 3, in the YAML).

Click the Knit button near the top (with the cute little ball of yarn next to it), and turn in the output (html file) to Canvas. 
