---
title: "Lab 3"
author: "Asha Yadav"
date: "October 16, 2019"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE,
                      warning = FALSE)
```

1. Run the following code to load the *tidyverse*.

```{r load-tidyverse}
library(tidyverse)
```

1. Run the following lines of code to view and read about the data we'll use 
for this lab.

```{r gsscat, eval = FALSE}
str(gss_cat)
View(gss_cat)
?gss_cat
```

2. Use `ggplot` to create a histogram of `tvhours`. The plot should:
  + Include a small amount of transparency so gridlines are visible under the 
    histogram.
  + Be a different color than the default
  + Include meaningful axis labels
  + Have a title and a subtitle

```{r question-2}
ggplot(gss_cat, aes(x = tvhours)) + 
  geom_histogram(alpha = 0.7, color = "red", fill = "cornflowerblue") + labs(x = "TV Hours per day", y = "Observations", title = "TV Hours per day", subtitle = "This is a subtitle")
```

3. Create a ridgeline density plot of `age` by `partyid`. Do you notice any strong
  differences by party affiliation?

```{r question-3}
library(ggridges)
ggplot(gss_cat, aes( age, partyid)) + geom_density_ridges(alpha = 0.7, color = "red", fill = "cornflowerblue")
# Strong democrat seem to have strong affiliation with high variability around the mean age of 50.
```

4. Modify the plot from Question 3 to produce separate ridgeline densities by `race`. Do 
  you see any patterns?

```{r question-4}
ggplot(gss_cat, aes( age, partyid)) + geom_density_ridges(aes(fill = factor(race), alpha = 0.7))
# Do not see any pattern.

```

5. Replicate the plots from Questions 3 and 4, but this time displaying the 
  relations with boxplots (i.e., one boxplot from Question 3 and 
  one boxplot for Question 4 in the chunk below). Which do you think displays the relation more 
  clearly? Why?
  Ans: Boxplot for Question 4 is better with more information and clear presentation. It shows outliers and race as well.
  
```{r label, options}
ggplot(gss_cat, aes(partyid, age)) + geom_boxplot(alpha = 0.7, color = "red", fill = "cornflowerblue") +  theme(axis.text.x = element_text(angle = 60, hjust = 1))
```
```{r}
ggplot(gss_cat, aes(partyid, age)) + geom_boxplot(aes(fill = factor(race), alpha = 0.7)) +  theme(axis.text.x = element_text(angle = 60, hjust = 1))
```

6. Run the following lines of code to get a correlation table as a data frame.

```{r corr-mat-df}
# You might need to install the ISLR package using the line below. 
# If so, uncomment out the line, install the pacakge, then comment the line again. 
# Your document won't knit if you do not comment the line below.

#install.packages(ISLR) 
#library(ISLR)
cor_tbl <- Hitters %>% 
  select_if(is.numeric) %>% 
  cor(use = "pairwise.complete.obs") %>% 
  as.data.frame() %>% 
  mutate(row = row.names(.)) %>% 
  gather(col, cor, -row) %>% 
  tbl_df()
```

Use the correlation table to create a heatmap using `geom_tile`. Change the 
colors to one of the four viridis palettes. What does the heatmap help you see?

```{r, heatmap}
ggplot(cor_tbl, aes(row, col, fill = cor)) + geom_tile() + scale_fill_viridis_c() + 
   theme(axis.text.x = element_text(angle = 60, hjust = 1))
# Lowest to highest correlation
```

7. The code below will produce a barplot. 

```{r barplot1}
ggplot(gss_cat, aes(rincome)) +
  geom_bar() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

Use the code below to create a new data set - `gss_cat_new` - that redefines the `rincome` variable so 
that *Not applicable* is displayed first, followed by *No answer*, *Don't know*, *Refused*, 
and then each income category from lowest to highest. Once you've redefined the factor
variable, re-produce the plot above by running the `ggplot` code at the bottom of the chunk. 
The plot should have *Not applicable* closest to the y-axis, then *No answer*, *Don't know*, etc.

```{r redfine-factor, echo = FALSE}
gss_cat_new <- gss_cat %>% 
  mutate(rincome = factor(rincome, levels = c("Not applicable", "No answer", "Don't know", "Regused", "Lr$1000", "$1000 to 2999", "$3000 t 3999", "$4000 to 4999", "$5000 to 5999", "$6000 to 6999", "$7000 to 7999", "$8000 to 9999", "$10000-14999", "$15000-19999", "$20000-24999", "$25000 or more")))

ggplot(gss_cat_new, aes(rincome)) +
  geom_bar() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```