---
title: "Lab 4"
author: "Asha Yadav"
date: "October 16, 2019"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

```{r setup, include=FALSE}
 knitr::opts_chunk$set(message = FALSE, 
                       warning = FALSE)
```


Install and load the package *Lahman*, which will give you access to the dataset *Teams*

```{r teams_echo, echo = TRUE}
#install.packages("Lahman")
library(Lahman)
library(tidyverse)
library(janitor)
Teams
```
* Produce a subset of the data (as a new object) that has the following characteristics:
    + Only one team (your choice)
    + data from 1980 to present (or as present as the dataset gets)
    + Includes 5 columns: name, yearID, W, L, R, RA

(The variables above correspond to the team name, the year, wins, losses, runs scored, and runs allowed)

* Make sure you select a team that is currently still around, or it probably
won't be interesting (see a list of current at [http://www.espn.com/mlb/teams](http://www.espn.com/mlb/teams)).

* Create a new variable corresponding to the winning percentage for the team 
  you chose over time

$$w_{pct} = \frac{wins}{wins + losses}$$

* Order by winning percentage: Least to greatest
* Order by winning percentage: greatest to least
* Compute the mean and standard deviation of winning percentage
* With the full dataset
  + compute the average and standard deviation of winning percentage for each team. 
  + Order by highest winning percentage
* Use the full data to reproduce the plot below

```{r answers}
subset_team <- Teams%>%
  select(name, yearID, W, L, R, RA)%>%
  filter(name == "Chicago Cubs" & yearID > "1980")%>%
  mutate(w_pct = W/(W+L))%>%
  arrange(w_pct)%>%
  arrange(desc(w_pct))%>%
  summarize(mean(w_pct), sd(w_pct))

Teams%>%
  select(name, yearID, W, L, R, RA)%>%
  mutate(w_pct = W/(W+L))%>%
  arrange(desc(w_pct))

  
  


```


```{r plot}
# Please put the code for the plot in this chunk.
Plot_subset <- Teams%>%
  select(name, yearID, W, L, R, RA)%>%
  filter(name %in% c("Detroit Tigers", "New York Yankees", "San Diego Padres"))%>%
  filter(yearID >= "1900") %>%
  mutate(w_pct = W/(W+L))%>%
  rename(Team = "name")%>%
  rename(year_id = "yearID")%>%
  arrange(year_id)

ggplot(Plot_subset, (aes(year_id, w_pct))) +
  geom_line(aes(color = Team))
  
  

  
  
  
  
  



```








