---
title: "Lab 5"
author: "Asha Yadav"
date: "October 23, 2019"
output: html_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE) 
```

1. Read in the following three datasets, using {here} and {rio}:
  * *sesame13.sav*
  * *star.csv*
  * *ais.xlsx*
  
*Hint*: For the *ais.xlsx* dataset look at the `skip` argument within the `rio` help 
  documentation.
  
```{r import}
library(tidyverse)
library(dplyr)
library(rio)
library(here)
here("DATA")

sesame <- import(here("DATA", "sesame13.sav"))
star <- import(here("DATA", "star.csv"))
ais <- import(here("DATA", "ais.xlsx"), skip = 15)


```

2. Using the ais data, compute the average red blood cell count and average bmi by sport. Output these data as SPSS and EXCEL files. 

```{r q2}

ais_spss <- ais %>%
  group_by(sport)%>% 
  summarise(mean_rcc = mean(rcc),
            mean_sport = mean(bmi))
  export(ais_spss, here("DATA", "ais_spss.csv"))
  
```

3. Use the sesame dataset to answer the following question: Was the average female age higher in schools or at home? 

```{r q3}
sesame %>%
  as_tibble()%>%
  characterize()%>%
  filter(sex == "Female")%>%
  group_by(setting)%>%
  summarise(mean_age = mean(age))
  
# Average female age is higher in school. 

```

4. First, how many rows and columns are in the *star* dataset? Next, remove outliers using a really poor method, just for practice, by eliminating students whose math (*tmathss*) scores were more than three standard deviations **above or below** the corresponding mean. How many rows are in the dataset now?

```{r q4}
dim(star)
# There are 5748 rows and 12 columns in the 'star' dataset.
star%>%
  filter(tmathss <= mean(tmathss) + 3*sd(tmathss) & 
  tmathss >=  mean(tmathss) - 3*sd(tmathss))%>%
  nrow()
# There are 5743 rows.
```
 
 
5. Use the star dataset to compute standardized math and reading scores (*tmathss* and *treadss*); i.e., for each of those variables, subtract the mean from each observation and divide by the standard deviation:$$x_s = \frac{x_i - \bar{X}}{sd(X)}$$
  + Check that your computations were correct by computing the mean and standard deviation of each variable (they should be 0 and 1). 
  + Compute the mean of the standardized variable for all *sex*/*frl* combinations (I'm asking you to extend what you know, we haven't talked explicitly about how to do this yet). 
  + What do you make of these findings? Do you see an effect by sex? An FRL effect? Is there evidence of an interaction (i.e., that the effect of FRL is greater for boys versus girls)?

```{r q5}

star_new <- star %>%
  mutate(stand_math = (tmathss - mean(tmathss)) / sd(tmathss), 
           stand_rdg = (treadss - mean(treadss)) / sd(treadss))

  star_new%>%
    summarize(mean(stand_math), sd(stand_math))
  
  star_new%>%
    summarize(mean(stand_rdg), sd(stand_rdg))
  
  star_new %>%
    group_by(sex, frl)%>%
    summarize(stand_math_mean = mean(stand_math), stand_math_sd = sd(stand_rdg))

#There is frl effect for sex for e.g stand_math_value is lower for boys and girls due to frl. Thus there is interaction between sex and frl.


```

